---
title: "Sven DS Final Report"
author: "Sven Jolivet"
date: "2023-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In the dynamic world of soccer, where every pass, sprint, and strategic move can influence the outcome of a game, the distance covered by a team throughout a match doesn't emerge as a pivotal metric. The modern game demands a a blend of skill, teamwork, and relentless physical effort. The distance traveled on the field is not merely a statistical figure but a reflection of the collective energy, endurance, and tactical acumen of the players.

## Questions 
Does running around the field more prompt you to shoot more?

Did running more effect who the tournaments winner was?

## Packages
```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
```

These packages were crucial in forming my conclusion and final thoughts.

## Unfiltered Data 
My data is from data.world and the author is Mikhail Zhilkin.

I was lucky enough to gather amazing data that was neat and didin't have any missing data. The author provided a plethora of various stats that I could use to make any conclusion I wanted to about the Euorpean Championship. 

https://data.world/cervus/uefa-euro-2020/workspace/file?filename=Match+team+statistics.csv

```{r}
TeamStats <- read_csv("TeamStats.csv", show_col_types = FALSE)
glimpse(TeamStats)
```

## Filtered Data 
Having 19,000 rows is pretty overwhelming and a lot of the data is not useful so I narrowed it down to key factors that helped me answer the questions and come to a conclusion. 

### Here are the key terms I used.

TeamName: Is the Name of the Team

StatsName: The specific stat that we are talking about 

MatchID: The specific match that was played

Value: The numerical value for StatsName

Distance: The distance a team ran 

Shots: The amount of shots a team took
```{r}
data <- TeamStats %>% select(TeamName, StatsName, Value, MatchID)
glimpse(data)
```

## Total Distance
I was able to find the total distance that each team ran during the tournament. 


```{r}
dist_cov <- data %>% 
  group_by(TeamName) %>% 
  filter(StatsName == "Distance covered (m)") %>% 
  mutate(Value = as.numeric(Value))

kable_dis <- dist_cov %>% group_by(TeamName) %>% 
  select(TeamName, Value) %>% 
  arrange(Value)

kable(kable_dis)
```

After reviewing the amounts for each team I was able to see that the Netherlands had the least distance covered.
This could be because of their play style. North European teams tend to pass the ball more than southern Europe. For instance, the tournament winners Italy, had the highest running distance. Spain and Italy have high distances probably because they dribble the ball more. 

### Average Distance
I made a variable named "avg_dist" which I used to compare the different teams

```{r}
avg_dist <- dist_cov %>% group_by(TeamName) %>% 
  select(Value,TeamName) %>% 
  summarise(mean(Value, na.rm = T))
```

## Comparing Distances


```{r}
Distance <- suppressWarnings(as.numeric(dist_cov$Distance))
```


Since it is a tournament not all teams play the same amount of games, so I needed to take the average of the games they did play in order to compare them. 
```{r}
names(avg_dist)[2] <- "Distance"
```

I needed to create the averages for the finalist vs 2 teams that didn't make it as far.

I used join to better help me filter and tidy the data.

```{r}
england_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "England") %>% 
  summarise(mean(Value))

```

```{r}
italy_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Italy") %>% 
  summarise(mean(Value))

wales_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Wales") %>% 
  summarise(mean(Value))

Netherlands_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Netherlands") %>% 
  summarise(mean(Value))

compare_dis <- full_join(italy_dis,england_dis)

compare_dis <- full_join(compare_dis, Netherlands_dis)

compare_dis <- full_join(compare_dis, wales_dis)

names(compare_dis)[2] <- "Value"

```

## Finalists 
As you can see the the finalist did run more than 2 teams that were eliminated earlier into the tournament. 

```{r}
dist_graph <- ggplot(compare_dis, aes(x=TeamName, y = Value, fill = TeamName))
dist_graph + geom_bar(stat = "identity")
```

## Total Attempts

```{r}

total_attempts <- data %>%
  group_by(TeamName) %>% 
  filter(StatsName == "Total Attempts") %>% 
  mutate(Value = as.numeric(Value)) %>% 
  arrange(Value)

kable_attempts <- total_attempts %>% group_by(TeamName) %>% 
  select(TeamName, Value)
  
  
kable(kable_attempts)
```

## Distance Vs Shots Attempted
Does running around the field more prompt you to shoot more?

I took the average shots attempted for each country and joined the average distance together.
In order to see if there is a trend between the two factors.

```{r}
avg_attempts <- total_attempts %>% group_by(TeamName) %>% 
  select(Value,TeamName) %>% 
  summarise(mean(Value, na.rm = T))

names(avg_attempts)[2] <- "Value"
kable(avg_attempts)
```

The highest shot average was Italy and the lowest was Finland

This shows how Italy had a very dominant tournament as they ended up winning it. 
 
```{r}
dist_attempts <- full_join(avg_attempts,avg_dist)
names(dist_attempts)[2] <- "Shots"
names(dist_attempts)[3] <- "Distance"

kable(dist_attempts)

```

## Chart

```{r}
avg_dist_plot <- ggplot(data = dist_attempts %>% select(Shots, Distance), mapping = aes(x = Shots, y = Distance))
avg_dist_plot + geom_point()

```

As shown in the graph there isn't a major trend in distance traveled compared to the amount of shots taken. There are major outliers that had high distance traveled but very few shots attempted. THe argument could be made that they are running a lot because they are defending much more. 


The team with the most shot attempts and average distance traveled ended up winning the tournament. 


## Summary
There are many important stats that can show why a team has more success than another team. My goal was to present a stat not so heavily represented and show how it can effect an outcome. I conclude that it isn't as conclusive as I had thought, but in the case of the 2020 European championship the team with the highest average distance ran won the tournament.