---
title: "Distance Traveled In Soccer"
author: "Sven Jolivet"
date: "2023-11-29"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

In the dynamic world of soccer, where every pass, sprint, and strategic move can influence the outcome of a game, the distance covered by a team throughout a match doesn't emerge as a pivotal metric. The modern game demands a a blend of skill, teamwork, and relentless physical effort. The distance traveled on the field is not merely a statistical figure but a reflection of the collective energy, endurance, and tactical acumen of the players.


## Questions 
Does running around the field more prompt you to shoot more?

Did running more effect who the tournaments winner was?

## Packages
```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
```


## Unfiltered Data 
My data is from data.world and the author is Mikhail Zhilkin.

https://data.world/cervus/uefa-euro-2020/workspace/file?filename=Match+team+statistics.csv

```{r}
TeamStats <- read_csv("TeamStats.csv", show_col_types = FALSE)
glimpse(TeamStats)

```

## Filtering
Having 19,000 rows is pretty overwhelming and a lot of the data is not useful so I narrowed it down to key factors that I will be looking for. 

TeamName: Is the Name of the Team

StatsName: The specific stat that we are talking about 

MatchID: The specific match that was played

Value: The numerical value for StatsName

Distance: The distance a team ran 

Shots: The amount of shots attempted
```{r}
data <- TeamStats %>% select(TeamName, StatsName, Value, MatchID)
glimpse(data)
```

## Total Distance
Finding the distance each team achieved during each of their games.

The netherlands had the lowest distance ran in a game and Italy had the highest.
```{r}
dist_cov <- data %>% 
  group_by(TeamName) %>% 
  filter(StatsName == "Distance covered (m)") %>% 
  mutate(Value = as.numeric(Value))

kable_dis <- dist_cov %>% group_by(TeamName) %>% 
  select(TeamName, Value) %>% 
  arrange(Value)

kable(kable_dis)


```



```{r}

avg_dist <- dist_cov %>% group_by(TeamName) %>% 
  select(Value,TeamName) %>% 
  summarise(mean(Value, na.rm = T))
```

## Comparing Distances


```{r}
Distance <- suppressWarnings(as.numeric(dist_cov$Distance))
```


Since it is a tournament not all teams play the same amount of games, so I needed to take the average of the games they did play in order to compare them. 
```{r}
names(avg_dist)[2] <- "Distance"
```


```{r}
england_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "England") %>% 
  summarise(mean(Value))

```

```{r}
italy_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Italy") %>% 
  summarise(mean(Value))

wales_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Wales") %>% 
  summarise(mean(Value))

Netherlands_dis <- dist_cov %>% group_by(TeamName) %>%
  filter(TeamName == "Netherlands") %>% 
  summarise(mean(Value))

compare_dis <- full_join(italy_dis,england_dis)

compare_dis <- full_join(compare_dis, Netherlands_dis)

compare_dis <- full_join(compare_dis, wales_dis)

names(compare_dis)[2] <- "Value"

```


## Finalists 
As you can see the the finalist did run more than 2 teams that were eleminated earlier into the tournament. 

```{r}
dist_graph <- ggplot(compare_dis, aes(x=TeamName, y = Value, fill = TeamName))
dist_graph + geom_bar(stat = "identity")
```

## Total Attempts

```{r}

total_attempts <- data %>%
  group_by(TeamName) %>% 
  filter(StatsName == "Total Attempts") %>% 
  mutate(Value = as.numeric(Value)) %>% 
  arrange(Value)

kable_attempts <- total_attempts %>% group_by(TeamName) %>% 
  select(TeamName, Value)
  
  
kable(kable_attempts)
```
## Distance Vs Shots Attempted
Does running around the field more prompt you to shoot more?

I took the average shots attempted for each country and joined the average distance together.
In order to see if there is a trend between the two factors.

```{r}
avg_attempts <- total_attempts %>% group_by(TeamName) %>% 
  select(Value,TeamName) %>% 
  summarise(mean(Value, na.rm = T))

names(avg_attempts)[2] <- "Value"
kable(avg_attempts)
```

The highest shot average was Italy and the lowest was Finland
 
```{r}
dist_attempts <- full_join(avg_attempts,avg_dist)
names(dist_attempts)[2] <- "Shots"
names(dist_attempts)[3] <- "Distance"

kable(dist_attempts)

```

## Chart

```{r}
avg_dist_plot <- ggplot(data = dist_attempts %>% select(Shots, Distance), mapping = aes(x = Shots, y = Distance))
avg_dist_plot + geom_point()

```


The team with the most shot attempts and average distance traveled ended up winning the tournament. 


## Summary
There are many important stats that can show why a team has more success than another team. My goal was to present a stat not so heavily represented and show how it can effect an outcome. I conclude that it isn't as conclusive as I had thought, but in the case of the 2020 European championship the team with the highest average distance ran won the tournament.

